---
title: "NHP: rationale for TPMA selections"
date: "last-modified"
date-format: D MMM YYYY HH:mm
author: "[Data Science, The Strategy Unit](mailto:mlcsu.su.datascience@nhs.net)"
format:
  html:
    grid:
      body-width: 1400px
execute:
  echo: false
  warning: false
resource_files:
- R/fetch.R
- R/params.R
- R/reasons.R
- reference/nhp-scheme-lookup.R
---

```{r}
#| label: attach-functions
#| include: FALSE

purrr::walk(
  file.path("R", c("fetch.R", "params.R", "reasons.R")),
  source
)
```

```{r}
#| label: read-reference-data

tpma_lookup <- fetch_tpma_lookup()

scheme_lookup <- readr::read_csv(
  "reference/nhp-scheme-lookup.csv",
  col_types = "c"
)
```

```{r}
#| label: get-tagged-runs

board <- pins::board_connect()
tagged_runs_meta <- board |> fetch_tagged_runs_meta()
tagged_runs_params <- board |> fetch_tagged_runs_params(runs_meta = tagged_runs_meta)
```

```{r}
#| label: extract-values

tagged_runs_reasons <- tagged_runs_params |>
  purrr::map(\(x) purrr::pluck(x, "reasons"))

tpma_reasons_df <- purrr::map(
  c("activity_avoidance", "efficiencies"), # separate list elements, need both
  \(x) extract_reasons(tagged_runs_reasons, x, tpma_lookup, scheme_lookup)
) |>
  purrr::list_rbind()

tpma_values_df <- extract_params(
  tagged_runs_params,
  tagged_runs_meta,
  tpma_lookup,
  scheme_lookup
) |>
  dplyr::filter(!is.na(`TPMA code`)) # discard mitigators not in lookup
```

```{r}
#| label: prepare-tables

# Final output tables

tpmas_by_scheme <- dplyr::left_join(
  tpma_values_df,
  tpma_reasons_df,
  by = dplyr::join_by(
    `TPMA code`,
    `TPMA name`,
    `TPMA type`,
    `Activity type`,
    Scheme
  )
) |>
  dplyr::rename(
    "Low (scheme)" = Low,
    "High (scheme)" = High,
    "Midpoint (scheme)" = Midpoint,
    "Range (scheme)" = Range
  )

tpma_values_summarised <- tpma_values_df |>
  dplyr::summarise(
    dplyr::across(
      c(Low, High, Midpoint, Range),
      mean
    ),
    .by = c(
      `TPMA code`,
      `TPMA name`,
      `TPMA type`,
      `Activity type`
    )
  ) |>
  dplyr::rename(
    "Low (mean)" = Low,
    "High (mean)" = High,
    "Midpoint (mean)" = Midpoint,
    "Range (mean)" = Range
  )
```

```{r}
#| label: combine-tables

tpmas_table <- tpmas_by_scheme |>
  dplyr::left_join(
    tpma_values_summarised,
    by = dplyr::join_by(
      `TPMA code`,
      `TPMA name`,
      `TPMA type`,
      `Activity type`
    )
  ) |>
  dplyr::mutate(
    `TPMA name` = paste0(`TPMA code`, ": ", `TPMA name`),
    `Years in horizon` = `Horizon year` - `Baseline year`,
    dplyr::across(dplyr::where(is.double), \(x) janitor::round_half_up(x, 3)),
    dplyr::across(
      c(`Activity type`, `TPMA type`, `TPMA name`, Scheme),
      factor  # allows for discrete filtering in DT::datatable()
    )
  ) |>
  dplyr::select(
    `Activity type`,
    `TPMA type`,
    `TPMA name`,
    Scheme,
    `Years in horizon`,
    tidyselect::starts_with("Midpoint"),
    tidyselect::starts_with("Range"),
    tidyselect::starts_with("Low"),
    tidyselect::starts_with("High"),
    Reason
  )
```

This table shows the types of potentially-mitigatable activity (TPMAs) selected by each of the New Hospital Programme (NHP) schemes and their rationale for doing so, if available.
The only scenarios shown are those that are labelled internally with a run stage of 'validation_report_ndg2', otherwise 'final_report_ndg2', i.e. the primary used in an outputs report.

The value for each  TPMA is summarised here as the midpoint (i.e. the difference between the low and high values). For context, the mean midpoint across all schemes with a final scenario is also shown.
The downloadable CSV also contains columns for the low, high and range values selected by the scheme and the mean of these values across all schemes that have run a 'validation' or 'final' scenario.

Please [message the Data Science team inbox](mailto:mlcsu.su.datascience@nhs.net) with questions or requests.

```{r}
#| label: interactive-table

tpmas_table |>
  DT::datatable(
    extensions = 'Buttons',
    options = list(
      dom = "Bftp",
      pageLength = 5,
      buttons = list(
        list(
          extend = "csv",
          filename = paste0(Sys.Date(), "_tpma-rationale"),
          text = "Download CSV"
        )
      ),
      columnDefs = list(
        list(
          visible = FALSE,
          targets = 7:12
        )
      )
    ),
    filter = "top",
    rownames = FALSE
  )
```
